Non-kissing and non-crossing complexes \\ for gentle algebras

### 2018-05-24 draft/accordionComplexSurfaces.tex
cellular

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{question}[theorem]{Question}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{assumption}[theorem]{Assumption}

@@

\begin{definition}
Let~$\surface$ be a surface with two sets~$M$ and~$M\dual$ of marked points so that~$M$ and~$M\dual$ alternate along the boundary of~$\surface$.
A dissection~$\dissection$ of~$(\surface, M)$ and a dissection~$\dissection\dual$ of~$(\surface, M\dual)$ are \defn{dual dissections} if there are pairs of mutually inverse bijections~$M\dual \leftrightarrow \faces(\dissection)$ and~$M \leftrightarrow \faces(\dissection\dual)$, denoted~$\dual$ in both directions, such that the dissection~$\dissection$ has an edge joining its vertices~$u,v \in M$ and separating its faces~$f,g \in  \faces(\dissection)$ if and only if~$\dissection\dual$ has an edge joining its vertices~$f\dual, g\dual \in M\dual$ and separating its faces~$u\dual, v\dual \in \faces(\dissection\dual)$.
\end{definition}

Note that contrarily to the usual conventions, the dual vertex of a face~$f$ is not always in the interior of the face.
More precisely, there are two situations:
\begin{itemize}
\item if a face~$f$ has no edge on the boundary of~$\surface$, its dual vertex~$f\dual$ lies in the interior~of~$f$,
\item if a face~$f$ has an edge on the boundary of~$\surface$, then it has exactly one such edge and its dual vertex~$f\dual$ lies on this edge.
\end{itemize}
In fact, the second point forces the following characterization of the dissections that admit a dual dissection.

\begin{proposition}
\label{prop:conditionsDualDissections}
For a dissection~$\dissection$ of a marked surface~$(\surface, M)$, there exists a set~$M\dual$ of points of~$\surface$ and a dissection~$\dissection\dual$ of~$(\surface, M\dual)$ such that~$\dissection$ and~$\dissection\dual$ are dual dissections if and only if each face of~$\dissection$ has at most one boundary edge.
Moreover, the set~$M\dual$ is unique.
\end{proposition}

\begin{proof}
\vincent{Todo}
\end{proof}

\begin{remark}
In Proposition~\ref{prop:conditionsDualDissections}, note that although the set~$M\dual$ of dual marked points is unique, the dual dissection~$\dissection\dual$ is not when~$\dissection$ is not cellular.
\end{remark}

\begin{example}
Two examples of dual dissections on a surface are represented in \fref{fig:dissections}.
\end{example}

\begin{definition}
Interior points, blossom points.
Half-edges.

@@

Let~$\surface$ be a surface with two sets~$M$ and~$M\dual$ of marked points and two dual dissections~$\dissection$ and~$\dissection\dual$.
Let~$I$ denote the interior points and $B$ the blossom points.

@@

\begin{proposition}
\label{prop:accordionsZigzags}
For two dual dissections~$\dissection$ and~$\dissection\dual$, the $\dissection$-accordions are precisely the $\dissection\dual$-zigzags and the $\dissection$-accordion complex coincides with the $\dissection\dual$-zigzag complex.
\end{proposition}

@@

\begin{definition}
\label{def:noncrossingComplex}
We call \defn{non-crossing complex} of the pair of dual dissections~$(\dissection, \dissection\dual)$ the simplicial complex~$\NCC = \AC = \ZC$.
\end{definition}

@@

\begin{remark}
A quick disclaimer about accordion complexes and zigzag complexes of arbitrary dissections.
As stated in Proposition~\ref{prop:conditionsDualDissections}, a dissection~$\dissection$ with a face containing more than one boundary edge does not admit a dual dissection~$\dissection\dual$.
This prevents to use the definitions given in this section.
Although its accordion complex (resp.~zigzag complex) could still be defined as in~\cite{MannevillePilaud-accordion} (resp.~\cite{GarverMcConville}) for the disk, the resulting complexes would be join of smaller accordion complexes (resp.~zigzag complexes) defined in this paper.
See \cite[Prop.~2.4]{MannevillePilaud-accordion} for a detailed statement in the case of the disk.
\end{remark}

@@

In this section, we show that non-kissing complexes and non-crossing complexes actually coincide.
Namely, any pair of dual dissections~$(\dissection, \dissection\dual)$ defines a gentle bound quiver~$\bar Q_{\dissection, \dissection\dual}$ such that the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.
Conversely, any gentle bound quiver~$\bar Q$ gives rise to a surface~$\surface_{\bar Q}$ equipped with a pair of dual dissections~$\dissection_{\bar Q}$ and~$\dissection\dual_{\bar Q}$ such that the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, \dissection\dual_{\bar Q}]$.
In fact, we show that the accordion, zigzag, and non-kissing complexes can all be pictured in a unique way on the surface.
Interestingly, via the correspondence between pairs of dual dissections and gentle bound quivers, the duality between the dissections~$\dissection$ and~$\dissection\dual$ translates to the Koszul duality between the gentle algebras of~$\bar Q$ and~$\bar Q\dual$.

@@

\begin{proposition}
\label{prop:dualityKoszul1}
For two dual dissections~$\dissection$ and~$\dissection\dual$ of an oriented surface, the quivers~$\bar Q_{\dissection}$ and~$\bar Q_{\dissection\dual}$ are Koszul duals.
\vincent{Here I mean that the gentle algebras are Koszul duals. We need to define that somewhere and discuss it on gentle quivers.}
\end{proposition}

@@

\begin{definition}
\label{def:surfaceQuiver}
The surface~$\surface(\bar Q)$ of a gentle bound quiver~$\bar Q = (Q,I)$ is obtained by glueing lozanges around arrows of~$Q$ following the rule given by the relations~$I$. The construction also automatically defines two dual dissections~$\dissection_{\bar Q}$ and~$\dissection\dual_{\bar Q}$ of the surface~$\surface_{\bar Q}$.
\end{definition}

@@

\begin{theorem}
\label{thm:complexesCoincide}
\begin{itemize}
\item For any gentle bound quiver~$\bar Q$, the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, \dissection\dual_{\bar Q}]$, therefore to the accordion complex~$\AC[\dissection_{\bar Q}]$ and to the zigzag complex~$\ZC[\dissection\dual_{\bar Q}]$.
\item For any pair of dual dissections~$\dissection, \dissection\dual$ of an oriented surface, the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.
\end{itemize}
\end{theorem}

>>>>

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{question}[theorem]{Question}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{convention}[theorem]{Convention}

@@

\begin{convention}
\textbf{All throughout the paper, all dissections are considered cellular.}
The case of non-cellular dissections will be discussed in Section~\ref{sec:nonCellularDissections}.
\end{convention}

\begin{definition}
Let~$\surface$ be a surface with two sets~$M$ and~$M\dual$ of marked points so that~$M$ and~$M\dual$ alternate along the boundary of~$\surface$.
A cellular dissection~$\dissection$ of~$(\surface, M)$ and a cellular dissection~$\dissection\dual$ of~$(\surface, M\dual)$ are \defn{dual cellular dissections} if there are pairs of mutually inverse bijections~$M\dual \leftrightarrow \faces(\dissection)$ and~$M \leftrightarrow \faces(\dissection\dual)$, denoted~$\dual$ in both directions, such that~$\dissection$ has an arc joining its vertices~$u,v \in M$ and separating its faces~$f,g \in  \faces(\dissection)$ if and only if~$\dissection\dual$ has an arc joining its vertices~$f\dual, g\dual \in M\dual$ and separating its faces~$u\dual, v\dual \in \faces(\dissection\dual)$.
\end{definition}

Note that contrarily to the usual conventions, the dual vertex of a face~$f$ is not always in the interior of the face.
More precisely, there are two situations:
\begin{itemize}
\item if a face~$f$ has no arc on the boundary of~$\surface$, its dual vertex~$f\dual$ lies in the interior~of~$f$,
\item if a face~$f$ has an arc on the boundary of~$\surface$, then it has exactly one such arc and its dual vertex~$f\dual$ lies on this arc.
\end{itemize}
In fact, the second point forces the following characterization of the cellular dissections that admit a dual cellular dissection.

\begin{proposition}
\label{prop:conditionsDualDissections}
For a cellular dissection~$\dissection$ of a marked surface~$(\surface, M)$, the following assertions are equivalent:
\begin{enumerate}[(i)]
\item there exists a set~$M\dual$ of points of~$\surface$ and a cellular dissection~$\dissection\dual$ of~$(\surface, M\dual)$ such that~$\dissection$ and~$\dissection\dual$ are dual cellular dissections,
\item each face of~$\dissection$ has at most one boundary arc.
\end{enumerate}
Moreover, the set~$M\dual$ and the cellular dissection~$\dissection\dual$ are both uniquely determined.
\end{proposition}

\begin{proof}
Assume first that~$\dissection$ and~$\dissection\dual$ are dual cellular dissections. If a face~$F$ of~$\dissection$ had two boundary arcs~$a, b$, then it would contain two points of~$M\dual$, one on~$a$ and one on~$b$.
Conversely, if each face of~$\dissection$ has at most one boundary arc, we construct the set of points~$M\dual$ with one point on each boundary arc of~$\surface$ and one point in the interior of each interior face of~$\dissection$. Clearly, $M\dual$ has one point~$f\dual$ in each face~$f$ of~$\dissection$. We then construct half-arcs of~$\dissection\dual$ in each face~$f$ of~$\dissection$ by joining the dual point~$f\dual$ to the middle of each boundary arc of~$f$ that is not a boundary arc of~$\surface$. For each arc~$a$ of~$\dissection$, the half-arcs incident to the middle of~$a$ in the two faces of~$\dissection$ containing~$a$ form the dual arc~$a\dual$ of~$a$. All these dual arcs do not cross (as the half-arcs do not cross in each face of~$\dissection$) and form the dual cellular dissection~$\dissection\dual$ of~$\dissection$.
More details can be found in~\cite{PlamondonSchroll}.
\end{proof}

%\begin{remark}
%In Proposition~\ref{prop:conditionsDualDissections}, note that although the set~$M\dual$ of dual marked points is unique, the dual dissection~$\dissection\dual$ is not when~$\dissection$ is not cellular.
%\end{remark}

\begin{example}
Two examples of dual cellular dissections on a surface are represented in \fref{fig:dissections}.
\end{example}

\begin{definition}
Interior points, blossom points.
Half-arcs.

@@

Let~$\surface$ be a surface with two sets~$M$ and~$M\dual$ of marked points and two dual cellular dissections~$\dissection$ and~$\dissection\dual$.
Let~$I$ denote the interior points and $B$ the blossom points.

@@

\begin{proposition}
\label{prop:accordionsZigzags}
For two dual cellular dissections~$\dissection$ and~$\dissection\dual$, the $\dissection$-accordions are precisely the $\dissection\dual$-zigzags and the $\dissection$-accordion complex coincides with the $\dissection\dual$-zigzag complex.
\end{proposition}

@@

\begin{definition}
\label{def:noncrossingComplex}
We call \defn{non-crossing complex} of the pair of dual cellular dissections~$(\dissection, \dissection\dual)$ the simplicial complex~$\NCC = \AC = \ZC$.
\end{definition}

@@

\begin{remark}
A quick disclaimer about accordion complexes and zigzag complexes of arbitrary dissections.
As stated in Proposition~\ref{prop:conditionsDualDissections}, a dissection~$\dissection$ with a face containing more than one boundary edge does not admit a dual cellular dissection~$\dissection\dual$.
This prevents to use the definitions given in this section.
Although its accordion complex (resp.~zigzag complex) could still be defined as in~\cite{MannevillePilaud-accordion} (resp.~\cite{GarverMcConville}) for the disk, the resulting complexes would be join of smaller accordion complexes (resp.~zigzag complexes) defined in this paper.
See \cite[Prop.~2.4]{MannevillePilaud-accordion} for a detailed statement in the case of the disk.
\end{remark}

@@

In this section, we show that non-kissing complexes and non-crossing complexes actually coincide.
Namely, any pair of dual cellular dissections~$(\dissection, \dissection\dual)$ defines a gentle bound quiver~$\bar Q_{\dissection, \dissection\dual}$ such that the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.
Conversely, any gentle bound quiver~$\bar Q$ gives rise to a surface~$\surface_{\bar Q}$ equipped with a pair of dual cellular dissections~$\dissection_{\bar Q}$ and~$\dissection\dual_{\bar Q}$ such that the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, \dissection\dual_{\bar Q}]$.
In fact, we show that the accordion, zigzag, and non-kissing complexes can all be pictured in a unique way on the surface.
Interestingly, via the correspondence between pairs of dual cellular dissections and gentle bound quivers, the duality between the dissections~$\dissection$ and~$\dissection\dual$ translates to the Koszul duality between the gentle algebras of~$\bar Q$ and~$\bar Q\dual$.

@@

\begin{proposition}
\label{prop:dualityKoszul1}
For two dual cellular dissections~$\dissection$ and~$\dissection\dual$ of an oriented surface, the quivers~$\bar Q_{\dissection}$ and~$\bar Q_{\dissection\dual}$ are Koszul duals.
\vincent{Here I mean that the gentle algebras are Koszul duals. We need to define that somewhere and discuss it on gentle quivers.}
\end{proposition}

@@

\begin{definition}
\label{def:surfaceQuiver}
The surface~$\surface(\bar Q)$ of a gentle bound quiver~$\bar Q = (Q,I)$ is obtained by glueing lozanges around arrows of~$Q$ following the rule given by the relations~$I$. The construction also automatically defines two dual cellular dissections~$\dissection_{\bar Q}$ and~$\dissection\dual_{\bar Q}$ of the surface~$\surface_{\bar Q}$.
\end{definition}

@@

\begin{theorem}
\label{thm:complexesCoincide}
\begin{itemize}
\item For any gentle bound quiver~$\bar Q$, the non-kissing complex~$\NKC$ is isomorphic to the non-crossing complex~$\NCC[\dissection_{\bar Q}, \dissection\dual_{\bar Q}]$, therefore to the accordion complex~$\AC[\dissection_{\bar Q}]$ and to the zigzag complex~$\ZC[\dissection\dual_{\bar Q}]$.
\item For any pair of dual cellular dissections~$\dissection, \dissection\dual$ of an oriented surface, the non-crossing complex~$\NCC$ is isomorphic to the non-kissing complex~$\NKC[\bar Q_{\dissection, \dissection\dual}]$.
\end{itemize}
\end{theorem}
